---
title: "GSS challange"
author: "MarenaO"
date: "Spring 2022 Semester"
output: html_document
---







```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,sex,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum,bigbang,god,postlife)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~race + cappun,data=mydata)
barchartGC(~race + cappun,data=mydata, type="percent")

```

According to the graphical data, I can assume that there is no difference between the variables. It seems like attitude towards capitol punishment is not independent of race due to the bar graphs looking similar to each other.

## Now with LESS data (more variables give less data due to NA's)

```{r}
dd2 <- mydata %>% group_by(bigbang,postlife) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=bigbang,y=count,fill=postlife))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=bigbang,y=prcnt*100,fill=postlife)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```
According to the numerical data, there seems to not be much of a difference between the races on their attitude towards capitol punishment. The ercentages for each race are around the same ballpark for them oppposing or favoring capitol punishment.

## Now with less data

```{r}
table2 <- xtabs(~bigbang + postlife, data=mydata)
rowPerc(table2)
colPerc(table2)
```


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
According to the inferential data, I can confirm that there is no statistically significant difference between the races attitudes towards capitol punishment. This can be confirmed by a p-vlaue of 0.5, which suggests that there is no statistically significant difference between them.

## Less data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
#fisherXY(mydata$race,mydata$cappun)
```
```{r}
fisherXY = function(x, y, confL = 0, printAll=FALSE){
  # Inputs: 
  #   x: factor variable, 2 levels
  #   y: factor variable, n levels
  # confL: confidence level, default = 0.95
  #        print only those tables, results if p-value < (1-confL)
  # printAll: overide the 'print only if p-value < (1-confL), and
  #           print all tables and fisher test results
  #
  # Constraint:
  #   x and y must be same length
  #
  # Outputs:
  #   All 2x2 Fisher exact tests
  
  if(confL<0 || confL> 1){
    print('error: confL must be between 0 and 1')
    return(999)
  }
  
  if(!is.factor(x)) x=as.factor(x)
  if(!is.factor(y)) y=as.factor(y)
  
  fisherTable = as.matrix(table(x, y))
  cat('\ncontingency table:\n')
  print(fisherTable)
  cat('\n***All 2X2 tests****\n\n')

  # all 2x2 tests
  levelsY = levels(y)
  levelsX = levels(x)
  
  m = length(levelsX)
  n = length(levelsY)
  
  confLevel = 0.95
  if (confL !=0 ) confLevel=confL
  
  for (i in 1:(m-1))
       for(j in ((i+1):m))
          for (k in 1:(n-1))
             for (l in ((k+1):n)) {
                
                 fisherIndexX = c(i,j)
                 fisherIndexY = c(k,l)
                 
                 fTempTable = as.matrix(fisherTable[fisherIndexX,
                                        fisherIndexY])
                 
                 fisherTest=fisher.test(fTempTable, conf.int = TRUE, 
                              conf.level = confLevel)
                 
                 if((fisherTest$p.value < (1-confLevel) ) || printAll){
                   
                       print(c(levelsX[i], levelsX[j]))
                       print(c(levelsY[k], levelsY[l]))
                       print(fTempTable)
                       print(fisherTest)
                       cat('\n')
                 }
                 
        }
}

```

```{r,eval=FALSE}
fisherXY(mydata$cappun,mydata$race)
```
 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The objective was to find if attitude towards capitol punishment is independent of race or not. The graphical results show there is no difference between the variables, since the bar graphs looked similar to each other. The numerical results showed that there was not much of a difference between the races on their attitude towards capitol punishment since the percentages were pretty similar to each other. Running the inferential results I concluded that there is no statistical significance between the data. This was confirmed using a Chi Squared test that gave a p-value of 0.5, which is high and can help see that there is no difference.

Is belief in postlife (postlife) independent of belief in the big bang theory (bigbang)?
If not, then whose view differs from whom??

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~bigbang + postlife,data=mydata)
barchartGC(~bigbang + postlife,data=mydata, type="percent")

```

According to the graphical data, we can assume that there is a difference between the variables. The chart on the right indicates that there are people who do not believe in the big bang theory, although most of the people believe in the afterlife. The chart on the left indicates that there are people who believe in the big bang theory and most of these people also believe in the afterlife.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~bigbang + postlife, data=mydata)
rowPerc(table1)
colPerc(table1)
```

According to the numerical data, it looks like there is a difference between the data. Although, it does not look like a dramatic difference.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
According to the inferential data, I can conclude that there is actually no statistically significant difference in the data. The Fisher exact test can confirm this. Since 1 is in the confidence interval, then that means that there is no difference in the data.